Ground state energy of quantum dots using the coupled cluster method

### 2018-04-16 project_2/article/article.tex
Add start of CC amplitude equations

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        v(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + W
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $W$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or de-exciting the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
    \end{align}
    for any excitation.

@@

    \subsection{Energy of the coupled cluster approximation}
        When we're going to compute the energy of a system using the coupled
        cluster approximation we would ideally want to find the expectation
        value of the energy using the coupled cluster wavefunction.
        \begin{align}
            \ecc = \bcc H\kcc.
        \end{align}
        As it turns out, this is an uncomfortable way of finding the energy
        as $T \neq T^{\dagger}$. Instead we will define what we call the
        \emph{similarity transformed Hamiltonian}. We plug the coupled
        cluster wavefunction into the Schrödinger equation.
        \begin{align}
            H\kcc = \ecc\kcc.
        \end{align}
        Next, we left multiply with the inverse of the cluster expansion,
        i.e.,
        \begin{align}
            e^{-T}H\kcc = e^{-T}\ecc\kcc
            = \ecc \kslat.
        \end{align}
        Projecting this equation on the reference state we get
        \begin{align}
            \ecc = \bslat e^{-T}H\kcc
            = \bslat e^{-T}He^{T}\kslat,
        \end{align}
        where in the latter inner-product we have located the similarity
        transformed Hamiltonian defined by
        \begin{align}
            \bar{H} \equiv e^{-T}He^{T}.
            \label{eq:similarity_transformed_hamiltonian}
        \end{align}

        To simplify the energy equation and the amplitude equations we use the
        normal ordered Hamiltonian.
        \begin{align}
            H = H_N + \bslat H\kslat.
        \end{align}
        The energy equation thus becomes
        \begin{align}
            \ecc &= \bslat\bar{H}\kslat
            = E_{\text{ref}} + \bslat e^{-T}H_N e^T\kslat,
        \end{align}
        where the reference energy is given by
        \begin{align}
            E_{\text{ref}} = \bslat H\kslat.
        \end{align}
        We now define the normal ordered similiarity transformed Hamiltonian as
        \begin{align}
            \bar{H}_N \equiv e^{-T}H_N e^T.
        \end{align}
        By expanding the exponentials of this Hamiltonian and recognizing the
        commutators we get the Baker-Campbell-Hausdorff expansion.
        \begin{align}
            \bar{H}_N
            &=
            H_N + \com{H_N}{T} + \frac{1}{2!}\com{\com{H_N}{T}}{T} + \dots.
        \end{align}
        From the connected cluster theorem we know that the only nonzero terms
        in the Baker-Campbell-Hausdorff expansion will be the terms where the
        normal ordered Hamiltonian has at least one contraction\footnote{In the
        Wick's theorem sense.} with every cluster operator on its right. This
        lets us write the expansion as
        \begin{align}
            \bar{H}_N
            &=
            H_N + \para{H_N T}_c + \frac{1}{2!}\para{H_N T^2}_c + \dots,
            \label{eq:sim_norm_hamiltionian_expansion}
        \end{align}
        where the subscript $c$ signifies that only contributions where at least
        one contraction between $H_N$ and $T$ has been performed will be
        included.

        \subsubsection{Coupled cluster doubles energy equation}
            Using the doubles approximation with the cluster operator $T_2$
            defined in \autoref{eq:T_2} the energy equation becomes
            \begin{align}
                \eccd
                &=
                E_{\text{ref}}
                + \bslat e^{-T_2}H_N e^{T_2}\kslat.
            \end{align}
            As the doubles cluster operator doubly excites the reference and
            using the expansion in \autoref{eq:sim_norm_hamiltionian_expansion}
            we see that we can write the energy equation as
            \begin{align}
                \eccd
                &=
                E_{\text{ref}}
                + \bslat H_N(1 + T_2)\kslat,
            \end{align}
            as the Hamiltonian is only able to de-excite one pair of single
            particle functions.

>>>>

\section{Theory}
    In this project we will study a system of $N$ interacting electrons. We will
    be looking at a Hamiltonian consisting of a one-body and a two-body part.
    The one-body part is given by
    \begin{align}
        h(\vf{r}_i)
        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
    \end{align}
    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
    The two-body part is the Coulomb interaction potential.
    \begin{align}
        v(\vf{r}_i, \vf{r}_j)
        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
    \end{align}
    We thus get the total Hamiltonian
    \begin{align}
        H &= h + W
        =
        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j),
    \end{align}
    where $h$ is the full one-body operator and $W$ the full two-body
    operator, i.e., over the entire system.  Working in a basis of $L$ single
    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
    determinant as
    \begin{align}
        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
    \end{align}
    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
    of the system. We call these single particle functions \emph{occupied} as
    they are contained in the Slater determinant.  We will denote the occupied
    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
    functions we can write this as
    \begin{align}
        \{\ket{p}\}_{p = 1}^L
        = \{\ket{i}\}_{i = 1}^N
        \cup
        \{\ket{a}\}_{a = N + 1}^L,
    \end{align}
    i.e., the general indexed states consists of both occupied and virtual
    states. Note that the single particle functions are orthonormal, i.e.,
    \begin{align}
        \braket{p}{q} = \delta_{pq}.
    \end{align}
    We can construct other Slater determinants in this basis by exciting
    or de-exciting the reference determinant. A general excitation is labeled
    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
    particle functions with indices $i, j, \dots$ from the reference and added
    $a, b, \dots$. Note that
    \begin{align}
        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
        \label{eq:excited_overlap}
    \end{align}
    for any excitation.

@@

    \subsection{Energy of the coupled cluster approximation}
        When we're going to compute the energy of a system using the coupled
        cluster approximation we would ideally want to find the expectation
        value of the energy using the coupled cluster wavefunction.
        \begin{align}
            \ecc = \bcc H\kcc.
        \end{align}
        As it turns out, this is an uncomfortable way of finding the energy
        as $T \neq T^{\dagger}$. Instead we will define what we call the
        \emph{similarity transformed Hamiltonian}. We plug the coupled
        cluster wavefunction into the Schrödinger equation.
        \begin{align}
            H\kcc = \ecc\kcc.
        \end{align}
        Next, we left multiply with the inverse of the cluster expansion,
        i.e.,
        \begin{align}
            e^{-T}H\kcc = e^{-T}\ecc\kcc
            = \ecc \kslat.
            \label{eq:clean_schrodinger}
        \end{align}
        Projecting this equation on the reference state we get
        \begin{align}
            \ecc = \bslat e^{-T}H\kcc
            = \bslat e^{-T}He^{T}\kslat,
        \end{align}
        where in the latter inner-product we have located the similarity
        transformed Hamiltonian defined by
        \begin{align}
            \bar{H} \equiv e^{-T}He^{T}.
            \label{eq:similarity_transformed_hamiltonian}
        \end{align}

        To simplify the energy equation and the amplitude equations we use the
        normal ordered Hamiltonian.
        \begin{align}
            H = H_N + \bslat H\kslat.
        \end{align}
        The energy equation thus becomes
        \begin{align}
            \ecc &= \bslat\bar{H}\kslat
            = E_0 + \bslat e^{-T}H_N e^T\kslat,
        \end{align}
        where the reference energy is given by
        \begin{align}
            E_0 = \bslat H\kslat.
        \end{align}
        We now define the normal ordered similiarity transformed Hamiltonian as
        \begin{align}
            \bar{H}_N \equiv e^{-T}H_N e^T.
        \end{align}
        By expanding the exponentials of this Hamiltonian and recognizing the
        commutators we get the Baker-Campbell-Hausdorff expansion.
        \begin{align}
            \bar{H}_N
            &=
            H_N + \com{H_N}{T} + \frac{1}{2!}\com{\com{H_N}{T}}{T} + \dots.
        \end{align}
        From the connected cluster theorem we know that the only nonzero terms
        in the Baker-Campbell-Hausdorff expansion will be the terms where the
        normal ordered Hamiltonian has at least one contraction\footnote{In the
        Wick's theorem sense.} with every cluster operator on its right. This
        lets us write the expansion as
        \begin{align}
            \bar{H}_N
            &=
            H_N + \para{H_N T}_c + \frac{1}{2!}\para{H_N T^2}_c + \dots,
            \label{eq:sim_norm_hamiltionian_expansion}
        \end{align}
        where the subscript $c$ signifies that only contributions where at least
        one contraction between $H_N$ and $T$ has been performed will be
        included.

        \subsubsection{Coupled cluster doubles energy equation}
            Using the doubles approximation with the cluster operator $T_2$
            defined in \autoref{eq:T_2} the energy equation becomes
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat e^{-T_2}H_N e^{T_2}\kslat.
            \end{align}
            As the doubles cluster operator doubly excites the reference and
            using the expansion in \autoref{eq:sim_norm_hamiltionian_expansion}
            we see that we can write the energy equation as
            \begin{align}
                \eccd
                &=
                E_0
                + \bslat H_N(1 + T_2)\kslat,
            \end{align}
            as the Hamiltonian is only able to de-excite one pair of single
            particle functions.

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
        \end{align}
        to find $t_{ij}^{ab}$.
